---
title: "Methylation Database Generation"
output:
  html_document:
    df_print: paged
---

Methylation data was collected by xxx and processed by xxx.
**Collaborators:** Shannon Dillon, Tony Iaculo, xxx

**Database approach being used due to large size of data.**

This is an R Markdown Notebook. When you execute code within the notebook, the results appear beneath the code.

## Setup

Import workflow adjusted from https://readxl.tidyverse.org/articles/readxl-workflows.html

### Load libraries
```{r message=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(DBI)
```

### I/O paths
```{r}
meth_path <- "/datasets/work/af-mlai-bio/work/Heat_Stress_Use-case/users/DIL041/window_read_means"
csv_names <- list.files(path = meth_path, pattern = "*window_mean_.*.txt", full.names = TRUE)

# Create a database connection
con <- DBI::dbConnect(RSQLite::SQLite(), dbname = "/datasets/work/af-mlai-bio/work/Heat_Stress_Use-case/users/DIL041/window_read_means/meth_db.sqlite")
```

### Animal ID mapping
Convert numeric IDs back to the animal IDs that were used for biochem data.
```{r}
#data for mapping between different animal IDs
map_path <- "/datasets/work/af-mlai-bio/work/Heat_Stress_Use-case/processed_data/Sample  key.xlsx" 

map_path %>%
  read_excel() %>% 
  write_csv("data/sample_key.csv") #generate csv for future quick view

animal_map <- read_excel(map_path, col_names = TRUE) %>%
  select(`window mean ID`, `Animal ID`) %>%
  distinct() %>%
  drop_na() %>%
  clean_names() %>%
  mutate(window_mean_id = as.character(window_mean_id),
         animal_id = as.character(animal_id))
animal_map
```

### Methylation data
Creating a function to add individual csv window files into the database. 
```{r warning=FALSE}
add_csv_db <- function(file, path, db) {
  file_name <- basename(file)
  print(file_name)
  parts <- str_split(file_name, "_", simplify = TRUE)
  treatment <- as.character(str_extract(parts[3], "\\d+"))
  chromosome <- as.character(str_extract(parts[4], "\\d+"))
  ID <- as.character(str_extract(parts[5], "\\d+"))

  # Get animal ID from animal_map
  animal_id <- animal_map$animal_id[match(ID, animal_map$window_mean_id)]
  
  positions <- read_csv(file.path(path, paste0("window_positions_chrm", chromosome, ".csv")), 
                        na = "nan", 
                        col_names = TRUE) %>%
    mutate(ID = animal_id) %>%
    rownames_to_column()

  d <- read_csv(file, col_types = "d", na = "nan", col_names = "methylation") %>%
    mutate(ID = animal_id) %>%
    mutate(chromosome = chromosome) %>%
    mutate(treatment = case_when(
      treatment == "3" ~ "pens",
      treatment == "2" ~ "hot",
      treatment == "1" ~ "pre"
    )) %>%
    rownames_to_column() %>%
    full_join(positions) %>%
    drop_na(methylation) %>%
    dplyr::select(-rowname)
  
  dbWriteTable(db, "meth_data", d, append = TRUE)
}
```
Iterate over the csv files and process them
```{r message = FALSE, results='markup'}
walk(csv_names, add_csv_db, path = meth_path, db = con)
```
Disconnect from the SQLite database
```{r message = FALSE}
dbDisconnect(con)
```
Further data analysis and processing in scripts biochem.Rmd and PCA.Rmd
