---
title: "Anova analysis on biochem data."
output:
  html_document:
    df_print: paged
---

Anova to quantify effect of treatment on biochemical data.

## Load libraries
```{r setup, include=FALSE}
library(tidyverse)
library(rstatix)
library(broom)
library(fs)
library("viridis")
library(htmltools)
library(knitr)
library(ggpubr)
library(emmeans)

knitr::opts_knit$set(root.dir = "/home/mce03b/R/methylation")
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE) 
```

## Read in biochem data
```{r, echo=FALSE}
file_path <- path("results", "biochem_clean.csv")

biochem <- read_csv(file_path) %>%
  mutate(treatment = as.factor(treatment),
         cohort = as.factor(cohort),
         room = as.factor(room),
         ID = as.factor(ID),
         days=as.integer(days))

# reorder treatment so it displays correctly in plots
biochem <- biochem %>%  mutate(treatment = fct_reorder(treatment, days, .fun='median'))
```    

## Summary statistics and visualisation

## Results {.tabset .tabset-fade}

### Statistics
```{r}
biochem_summary <- biochem %>%
  group_by(treatment) %>%
  get_summary_stats(sodium, type = "mean_sd")

# Print the summary table
print(biochem_summary)
```

### Room
```{r}
ggplot(biochem, aes(x = treatment, y = sodium, color = room)) +
  geom_jitter(height = .2, width = .2, alpha = .7) +
  scale_color_viridis(discrete = TRUE, option = "D") +
  theme_bw()
```

### Cohort
```{r}
## Second
ggplot(biochem, aes(x = treatment, y = sodium, color = cohort)) +
  geom_jitter(height = .2, width = .2) +
  scale_color_viridis(discrete = TRUE, option = "D") +
  theme_bw()
```

### Boxplot
```{r}
ggboxplot(biochem, x = "treatment", y = "sodium") + 
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE) 
```

## Check assumptions
```{r}
model  <- lm(sodium ~ treatment, data = biochem)
```
## Results {.tabset .tabset-fade}

### qqplot
Testing for normality
```{r}
ggqqplot(residuals(model)) #normality
print("Shapiro test")
shapiro_test(residuals(model)) #normality
```

### variance
Testing for homogeneity of variance
```{r}
plot(model, 1) #homogeneity of variance
print("Levene test")
biochem %>% levene_test(sodium ~ treatment) #homogeneity of variance
```

## One-way anova
Summary plot
```{r}
residuals <- biochem %>% anova_test(sodium ~ treatment)
pairwise_comp <- biochem %>% tukey_hsd(sodium ~ treatment)

pairwise_plot <- pairwise_comp %>% add_xy_position(x = "treatment")
ggboxplot(biochem, x = "treatment", y = "sodium") +
  stat_pvalue_manual(pairwise_plot, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(residuals, detailed = TRUE),
    caption = get_pwc_label(pairwise_plot)
    )
```

## Results {.tabset .tabset-fade}

### Residuals
One-way anova results
```{r}
print(residuals)
```

### Pairwise
Pairwise comparison (tukey)
```{r}
print(pairwise_comp)
```


## Two-way analysis

## Summary statistics

Grouped by cohort:

## Results {.tabset .tabset-fade}

### Stats 
```{r}
biochem %>%
  group_by(treatment, cohort) %>%
  get_summary_stats(sodium, type = "mean_sd")
```

### Plot
```{r}
ggboxplot(biochem, x = "treatment", y = "sodium", color='cohort') +
  scale_color_viridis(discrete = TRUE, option = "D") 
```

Grouped by room:

## Results {.tabset .tabset-fade}

### Stats
```{r}
biochem %>%
  group_by(treatment, room) %>%
  get_summary_stats(sodium, type = "mean_sd")
```

### Plot
```{r}
ggboxplot(biochem, x = "treatment", y = "sodium", color='room') +
  scale_color_viridis(discrete = TRUE, option = "D") 
```

## Check assumptions
```{r}
model_cohort <- lm(sodium ~ treatment*cohort, data = biochem)
model_room <- lm(sodium ~ treatment*room, data = biochem)
```

## Results {.tabset .tabset-fade}

### qqplot_C
Testing for normality in treatment*cohort model
```{r}
ggqqplot(residuals(model_cohort)) + ggtitle("Normality Plot")
print('Shapiro test')
shapiro_test(residuals(model_cohort)) #normality
```

### variance_C
Testing homogeneity of variance in treatment*cohort model
```{r}
plot(model, 1)
print("Levene test")
biochem %>% levene_test(sodium ~ treatment*cohort)
```

### qqplot_R
Testing for normality in treatment*room model
```{r}
ggqqplot(residuals(model_room)) + ggtitle("Normality Plot")
print("Shapiro test")
shapiro_test(residuals(model_room)) #normality
```

### variance_R
Testing homogeneity of variance in treatment*cohort model
```{r}
plot(model, 1)
print("Levene test")
biochem %>% levene_test(sodium ~ treatment*room)
```


## Anova (two-way)

Model treatment*cohort: 
```{r}
residuals_cohort <- biochem %>% anova_test(sodium ~ treatment*cohort)

pwc_cohort <- biochem %>% 
  group_by(cohort) %>%
  emmeans_test(sodium ~ treatment, p.adjust.method = "bonferroni") 
```

## Results {.tabset .tabset-fade}

### Plot
```{r}
pwc_cohort <- pwc_cohort %>% add_xy_position(x = "cohort")
ggboxplot(biochem, x = "cohort", y = "sodium", color='treatment') +
  scale_color_viridis(discrete = TRUE, option = "D") +  
  stat_pvalue_manual(pwc_cohort, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(residuals_cohort, detailed = TRUE),
    caption = get_pwc_label(pwc_cohort)
    )
```

### residuals
```{r}
print(residuals_cohort)
```

### Pairwise
```{r}
print(pwc_cohort)
```

Model treatment*room: 

```{r}
residuals_room <- biochem %>% anova_test(sodium ~ treatment*room)
pwc_room <- biochem %>% 
  group_by(room) %>%
  emmeans_test(sodium ~ treatment, p.adjust.method = "bonferroni") 
```

## Results {.tabset .tabset-fade}

### Plot
```{r}
pwc_room <- pwc_room %>% add_xy_position(x = "room")
ggboxplot(biochem, x = "room", y = "sodium", color='treatment') +
  scale_color_viridis(discrete = TRUE, option = "D") +  
  stat_pvalue_manual(pwc_room, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(residuals_room, detailed = TRUE),
    caption = get_pwc_label(pwc_room)
    )
```

### Residuals
```{r}
print(residuals_room)
```

### Pairwise:
```{r}
print(pwc_room)
```

