---
title: "Anova analysis on biochem data."
output: html_notebook
---

Anova to quantify effect of treatment on biochemical data, following on from visualisation.

### Load libraries
```{r setup, include=FALSE}
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(broom)
library(fs)
knitr::opts_knit$set(root.dir = "/home/mce03b/R/methylation")


```

### Read in biochem data
```{r}
file_path <- path("results", "biochem_clean.csv")
biochem_tibble <- read_csv(file_path) %>%
  mutate(treatment = as.factor(treatment),
         cohort = as.factor(cohort),
         room = as.factor(room),
         ID = as.factor(ID),
         days=as.integer(days))
        
```
```{r}
model <- aov(sodium ~ treatment + cohort + room, data = biochem_tibble)
summary(one_way)
```
```{r}
par(mfrow=c(2,2))
plot(model)
par(mfrow=c(1,1))
```
```{r}
tukey.two.way<-TukeyHSD(model)

tukey.two.way
```
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
